---
title: "a04"
editor: visual
author: Jorge Coronado
---

```{r}
library(readr)
library(dplyr)
library(here)
library(lubridate)
```

# Exercise 1: Practice Joins

## Read in the data

```{r}
bird_observations<- read.csv(here::here("data", "raw","52_pp52_birds_1.csv"))
surveys <- read.csv(here::here("data", "raw","52_pp52_surveys_1.csv"))
sites <- read.csv(here::here("data", "raw","52_pp52_sites_1.csv"))
taxalist <- read.csv(here::here("data", "raw","52_pp52_taxalist_1.csv"))
```

## Get familiar with the data

```{r}
dim(bird_observations)
```

```{r}
head(bird_observations)
```

```{r}
glimpse(bird_observations)
```

```{r}
summary(bird_observations)
```

```{r}
names(bird_observations)
```

```{r}
#unique(bird_observations) # I made this a comment to avoid render a full table
```

## Create a subset of bird_observations

```{r}
birds_subset <- bird_observations %>% 
  filter(species_id %in% c("BHCO", "RWBL")) %>%
  filter(site_id %in% c("LI-W", "NU-C"))

head(birds_subset)
```

## Use left_join() to merge birds_subset and sites tables

```{r}
birds_left <- birds_subset %>% 
  left_join(y = sites, by = "site_id")

head(birds_left)
```

## Use full_join() to merge birds_subset and sites tables

```{r}
birds_full <- full_join(x = birds_subset, y = sites, by = "site_id")

head(birds_full)
```

## Use inner_join() to merge birds_subset and taxalist data

```{r}
birds_inner <- inner_join(x = birds_subset, y = taxalist, by = "species_id")

head(birds_inner)
```

# Exercise 2: Practice Wrangling & Joining Data

## Wrangle bird_observations data and merge the data with all the other tables (sites, surveys, and taxalist)

```{r}
bird_observations <- bird_observations %>% 
  rename(bird_obs_notes = notes)

head(bird_observations)
```

```{r}
bird_obs_subset <- bird_observations %>% 
  full_join(y = taxalist, by = "species_id") %>% 
  full_join(y = sites, by = "site_id") %>% 
  full_join(y = surveys, by = c("site_id", "survey_id")) %>% 
  select(survey_date, common_name, park_name, bird_count, observer)

head(bird_obs_subset)
```

## Explore observer data and fix the values within this column so that all values are in the same format.

```{r}
unique(bird_obs_subset$observer)
```

```{r}
bird_obs_subset <- bird_obs_subset %>% 
  mutate(observer = if_else(condition = observer == "Josh Burns",
                            true = "J. Burns",
                            false = observer))

head(bird_obs_subset)
```

```{r}
unique(bird_obs_subset$observer)
```

# Assignment a04

```{r}
bird_obs_subset <- bird_obs_subset %>%
  mutate(
    survey_month = month(survey_date),
    survey_month = as.factor(survey_month)
  )

head(bird_obs_subset)
```

```{r}
bird_obs_subset <- bird_obs_subset %>% 
  relocate(survey_month, .after = survey_date)

head(bird_obs_subset)
```

```{r}
bird_obs_subset <- bird_obs_subset %>% 
  filter(park_name %in% c("Lindo", "Orme", "Palomino", "Sonrisa"))

unique(bird_obs_subset$park_name)
```

```{r}
birds_by_park_month <- bird_obs_subset %>% 
  group_by(park_name, survey_month) %>% 
  summarize(total_birds = sum(bird_count, na.rm = TRUE))

head(birds_by_park_month)
```

## Summary

The final data frame shows the total number or birds observed by park and survey month, which is the outcome I expected after using group_by() and summarize().

The results are informative because they allow us to compare bird observations across different parks and months.

The process could be improved by organizing the steps more clearly.
