---
title: "a05"
author: "Jorge Coronado"
---

# Introduction

The goal of this analysis is to explore visitation patterns in different restoration areas of the Sacramento - San Joaquin Delta. By looking at the number and type of visitors, we can better understand how people use these environments. This information is useful for managing these areas and making decisions about conservation and public access.

# Analysis

## Loading libraries

```{r}
library(readr)
library(here)
library(dplyr)
library(tidyr)
library(forcats) # makes working with factors easier
library(ggplot2)
library(leaflet) # interactive maps
library(DT) # interactive tables
library(scales) # scale functions for visualization
library(janitor) # expedite cleaning and exploring data
library(viridis) # colorblind friendly color palette
```

## Read in the data

```{r}
delta_visits_raw<- read.csv(here::here("data", "raw","Socioecological_monitoring_data.csv"))
```

## Getting the data ready

```{r}
delta_visits <- delta_visits_raw %>% 
  janitor::clean_names()

colnames(delta_visits)
```
```{r}
visits_long <- delta_visits %>% 
  pivot_longer(col = c(sm_boat, med_boat, lrg_boat, bank_angler, scientist, cars),
               names_to = "visitor_type",
               values_to = "quantity") %>% 
  rename(restore_loc = eco_restore_approximate_location) %>% 
  select(-notes)

daily_visits_loc <- visits_long %>% 
  group_by(restore_loc, date, visitor_type) %>% 
  summarise(daily_visits = sum(quantity))

daily_visits_totals <- daily_visits_loc %>% 
  group_by(restore_loc) %>% 
  mutate(n = sum(daily_visits)) %>% 
  ungroup() %>% 
  mutate(restore_loc = fct_reorder(restore_loc, desc(n)))

locations <- visits_long %>% 
  distinct(restore_loc, .keep_all = TRUE) %>% 
  select(restore_loc, latitude, longitude)
```

## Plot

```{r}
# create a custom theme for consistent and recurring styling
my_theme <- theme_bw(base_size = 14) +
  theme(legend.position = "bottom", # move legend to bottom
        axis.ticks.y = element_blank()) # remove y-axis ticks

# create bar plot with total visits by location and visitor type
ggplot(data = daily_visits_totals, # use summarized visits dataset
       aes(x = daily_visits, y = restore_loc,
           fill = visitor_type)) + # map visits, location, and visitor type
  geom_col() + # create bar chart
  scale_fill_viridis_d() + # apply colorblind-friendly palette
  labs(x = "Number of Visits", # label for x-axis
       y = "Restoration Location", # label for y-axis
       fill = "Type of Visitor", # legend title
       title = "Total Number of Visits to Delta Restoration Areas by visitor type", # main title
       subtitle = "Sum of all visits during study period") + # subtitle
  scale_x_continuous(breaks = seq(0, 120, 20), expand = c(0, 0)) + # control x-axis scale
  my_theme # apply my custom theme
```

## Map

```{r}
# create an interactive map using Leaflet
leaflet(locations) %>% # load dataset with location coordinates
  addProviderTiles(providers$Esri.WorldImagery) %>% # add satellite imagery background
  addProviderTiles(providers$CartoDB.PositronOnlyLabels) %>% # add labels on top
  addCircleMarkers(
    lng = ~ longitude, # longitude for marker position
    lat = ~ latitude, # latitude for marker position
    popup = ~ restore_loc, # show location name when clicked
    radius = 5, # size of markers
    fillColor = "salmon", # marker color
    fillOpacity = 1, # marker transparency
    stroke = TRUE, # draw border around markers
    weight = 0.5, # border thickness
    color = "white", # border color
    opacity = 1) #border opacity
```
# Discussion

The bar plot was chosen because it clearly shows the total number of visits for each restoration location and allows easy comparison between different visitor types. The use of color also helps distinguish categories quickly.

The map was selected because it provides a spatial perspective of the restoration areas. It helps the viewer understand where each location is and how they are distributed across the Delta. The interactive features, such as popups, make it more engaging and informative.
